Algorithm using wpd values for clustering 100 households

### 2021-08-18 supplementary/algorithm1/real-data/cluster4-iter100.Rmd
cleaning folders

## nobs per combination is 4

```{r, echo = TRUE}
ntimes_val = 4
```

```{r}
change_index <- function(data){

y <- y %>%
  mutate(time = row_number())

return(y)
}

endbreaks<- nrow(sim_panel_null)

p1 <- change_index(sim_panel_null) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p2 <- change_index(sim_panel_varf) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p3 <- change_index(sim_panel_varx) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue")+
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

p4 <- change_index(sim_panel_varall) %>%
  ggplot(aes(x = time,
             y = sim_data)) +
  geom_line() +
  scale_x_continuous(breaks = seq(1, endbreaks, nfacet_val*nx_val))+
  theme_bw() +
  geom_point(alpha = 0.5, color = "blue") +
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
  panel.grid.minor.x =  element_blank())+
  ylab("simulated response")+
  xlab("index")

t1 <- sim_panel_varall %>% ungroup() %>%  select (-c(1, 2))

# %>% 
#   kable(format = "latex", caption = "Simulated data for categories format")

t2 <- change_index(sim_panel_varall) %>% select (-c(1, 2))

# %>% 
#   kable(format = "latex",
#         caption = "Simulated data for time series format")

kable(list(t1, t2), format = "latex", caption = "Simulated data across combination of categories (left) and manipulated time series format (right)") %>% 
  row_spec(1:24,  background =  "#d3d3d3")

# kable(t1) %>%
#   kable_styling(full_width = FALSE, position = "float_left")
# kable(t2) %>%
#   kable_styling(full_width = FALSE, position = "left")
```

```{r plot4}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
```

## nobs per combination is 10

```{r nobs10, echo = TRUE}
ntimes_val = 10 # nobs per combination 
```

```{r plot10}
(p1 + p_null)/( p2 + p_varf)/(p3 + p_varx)/(p4 + p_varall)
```

@@

```{r, echo=TRUE}
sample_seed <-  seq(10, 100, 10)
```

@@

```{r data-make}
data_null <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_null_normal) %>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_null")

data_varf <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varf_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varf")

data_varx <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varx_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varx")

data_varall <- map(seq_len(sample_seed), function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varall_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varall")

@@

ggplot(heatmap_final,
       aes(x= reorder(value, data_type), y = as.factor(D))) +
  geom_tile(aes(fill = sim_data_quantile)) +
  scale_fill_distiller(palette = "YlGnBu")+ 
  facet_wrap(~category, ncol = 1) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Deciles") +
  xlab("data sets")

>>>>

@@

```{r, echo=TRUE}
sample_seed <-  seq(1, 100, 1)
```

@@

```{r data-make}
data_null <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_null_normal) %>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_null")

data_varf <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varf_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varf")

data_varx <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varx_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varx")

data_varall <- map(sample_seed, function(seed){
  set.seed(sample_seed[seed])
  change_index_data(5, 2, 3, sim_varall_normal)%>%
    compute_quantiles(quantile_prob_val)
}) %>%
  bind_rows(.id = "seed_id") %>%
  mutate(data_type = "data_varall")

@@

ggplot(heatmap_final,
       aes(x= reorder(value, data_type), y = as.factor(D))) +
  geom_tile(aes(fill = sim_data_quantile)) +
  scale_fill_distiller(palette = "YlGnBu", direction = 1)+ 
  facet_wrap(~category, ncol = 1) +
  theme(axis.text.x = element_text(angle = 90)) +
  ylab("Deciles") +
  xlab("data sets")

### 2021-08-18 supplementary/algorithm1/real-data/realdata-cust10-noiter.Rmd
cleaning folders

```{r}

elec_ts <- read_rds("../data/elec_ts.rds")

set.seed(12345)

sm_100 <- elec_ts %>%
  as_tibble() %>% 
  distinct(customer_id) %>%
  slice_sample(n = 100)

sm <- elec_ts %>% 
  filter(customer_id %in% sm_100$customer_id) %>% 
  ungroup()

sm_scaled <- sm %>%
  group_by(customer_id) %>% 
  dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
  ungroup()
```

>>>>

```{r}
sm <- gravitas::smart_meter10 
sm_scaled <- sm %>%
  group_by(customer_id) %>% 
  dplyr::mutate(general_supply_kwh = scale(general_supply_kwh)) %>% 
  ungroup()
```

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 10 households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

sm_scaled %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1, 
             strip.position = "right") +
  theme_bw() +
  xlab("time") +
  ylab("demand (in Kwh)")
```

```{r raw-sub, fig.cap="The raw plots for ten customers shown for Sept 2013-Oct 2013. The data is zoomed in and the y-scales made free to emphasize weekly, hourly or any behavior/patterns that they might have. There is some daily and weekly pattern in most households."}
sm %>% 
  filter(year(reading_datetime) == 2013 &
           month(reading_datetime) %in% c(9, 10)) %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, ncol = 1, scales = "free_y", 
             strip.position = "right") +
  theme_bw()+
  ylab("demand (in Kwh)") +
  theme(panel.grid.major.x = element_blank()) +
  scale_x_datetime("Date", date_labels = "%b %d",
                   breaks = "1 week",
                   date_minor_breaks = "1 day")  + 
  theme(panel.grid.major.x =  element_line(colour = "#A9A9A9"),
        panel.grid.minor.x =  element_line(colour = "#D3D3D3"))

### 2021-08-18 supplementary/algorithm1/real-data/realdata-cust2-iter5.Rmd
cleaning folders

```{r}
sm <- gravitas::smart_meter10 %>%
  filter(customer_id %in% c("10006704", "10017936", "10006414", "10018250", "10018064"))
```

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

@@

```{r raw-iter, fig.cap = "Five iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data. The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
all_data_clust %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("time") +
  ylab("demand (in Kwh)")
```

@@

_k = 3_
```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 3, border = 2:5)
```

_k = 4_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 4, border = 2:5)
```

_k = 5_

```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 5, border = 2:6)
```

>>>>

```{r}
sm <- gravitas::smart_meter10 %>%
  filter(customer_id %in% c("10006704", "10017936"))
```

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for 5 households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}

library(lubridate)

@@

```{r raw-iter, fig.cap = "Five iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data. The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
all_data_clust %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id, nrow = 2) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("time") +
  ylab("demand (in Kwh)")
```

@@

### 2021-08-18 supplementary/algorithm1/real-data/realdata-cust2.Rmd
cleaning folders

---
title: "Clustering real data with three customers and choosing optimal number of clusters"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false    
---

@@

four households are considered for the following analysis from the SGSC data set.
The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`. They consist of half-hourly data from 2012-2014. Figure \ref{fig:raw-all} shows how the raw time plots for these four households look like. Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).

```{r}
sm <- gravitas::smart_meter10 %>%
  filter(customer_id %in% c("10006704", "10017936", "10018250"))
```

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for four households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

@@

The distribution of demand across different hours of the day observed for these four households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}). The following characterization is done to validate results of the clustering approach later on.

```{r hod-box, fig.cap="The raw plots for four households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
sm %>% 
  create_gran("hour_day") %>% 
  ggplot(aes(x = hour_day, y = general_supply_kwh)) + 
  geom_boxplot() + 
  facet_wrap(~customer_id, ncol = 1, scales = "free_y") 
```

@@

```{r hod-deciles, fig.cap = "The deciles plots for four households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ. Except for the 90th decile, the deciles for the 4th household look pretty flat."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)")
```

@@

```{r raw-iter, fig.cap = "four iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data. The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
all_data_clust %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("time") +
  ylab("demand (in Kwh)")
```

@@

_k = 3_
```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 3, border = 2:5)
```

>>>>

---
title: "Clustering real data with two customers and choosing optimal number of clusters"
output:
  bookdown::pdf_book:
    #base_format: rticles::asa_article
    fig_height: 5
    fig_width: 8
    fig_caption: yes
    dev: "pdf"
    keep_tex: yes
    toc: false    
---

@@

two households are considered for the following analysis from the SGSC data set.
The data sets contain three columns `customer_id`, `reading_datetime` and `general_supply_kwh`. They consist of half-hourly data from 2012-2014. Figure \ref{fig:raw-all} shows how the raw time plots for these two households look like. Since all the dataset looks squeezed on this linear time scale, Figure \ref{fig:raw-sub} shows how the raw plot looks for 2 months (Sept 2013 and Oct 2013).

```{r}
sm <- gravitas::smart_meter10 %>%
  filter(customer_id %in% c("10006704", "10017936"))
```

```{r raw-all, fig.cap="The raw time plots for demand shown for the entire observation period between 2012-2014 for two households (facets). The data is squeezed stopping us from seeing any behavioral patterns."}
library(lubridate)

@@

The distribution of demand across different hours of the day observed for these two households through heatmaps (Figure \ref{fig:hod-deciles}) and quantile plots (Figure \ref{fig:hod-heatmap}). The following characterization is done to validate results of the clustering approach later on.

```{r hod-box, fig.cap="The raw plots for two households shown for Sept 2013-Oct 2013. The data is zoomed in to emphasize weekly, hourly or any behavior/patterns that they might have.", eval = FALSE}
sm %>% 
  create_gran("hour_day") %>% 
  ggplot(aes(x = hour_day, y = general_supply_kwh)) + 
  geom_boxplot() + 
  facet_wrap(~customer_id, ncol = 1, scales = "free_y") 
```

@@

```{r hod-deciles, fig.cap = "The deciles plots for two households are shown for entire observation period across hours of the day. The y-scales are made free to allow us to see daily patterns for each households. The deciles for all households tend to show an increase in the morning and evening hours, although the hours differ. Except for the 90th decile, the deciles for the 4th household look pretty flat."}
data_heatmap %>% 
  ggplot() + geom_line(aes(x=category, 
                           y = quantile_val, 
                           color = quantile_prob)) + facet_wrap(~customer_serial_id, ncol = 1, scales="free_y") + 
  scale_color_brewer(palette = "Blues") + 
  #scale_y_log10() +
  theme_bw() +
  xlab("hours of the day") +
  ylab("demand (in Kwh)")
```

@@

```{r raw-iter, fig.cap = "two iterations for each customer are considered by adding random noise $(N(0, \\sigma^2)$, where $\\sigma^2$ is very small relative to the the variance of the data. The raw plots for all simulated dataset is shown to make sure they look similar to the structure."}
all_data_clust %>% 
  ggplot(aes(x = reading_datetime, y = general_supply_kwh)) +
  geom_line() +
  facet_wrap(~customer_id) +
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  xlab("time") +
  ylab("demand (in Kwh)")
```

@@

_k = 2_
```{r}
plot(hc, cex = 0.6)
rect.hclust(hc, k = 2, border = 2:5)
```

### 2021-08-18 supplementary/algorithm2/real-data/algo2-elec100-diffscale.Rmd
cleaning folders

```{r}
# ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
# ggpairs_nocolor

@@

# ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)

ggpairs_nocolor <- ggscatmat(data_pcp_ge2, columns=3:7) +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:7, color = "group") +
  scale_colour_brewer(palette="Set2") + theme(legend.position = "bottom")

ggpairs_nocolor
```

```{r}
ggpairs_color
```

```{r}
# 
# data_order <- data %>% 
#   group_by(comb) %>% 
#   summarise(m = mean(wpd)) %>% 
#   arrange(desc(m))
#   
# data_median = data %>% 
#   group_by(group, comb) %>% 
#   summarize(wpd_med = median(wpd))
# 
# data_pcp <- data_median %>% 
# mutate(group = as.character(group)) %>% 
#   left_join(data_order) %>% 
#   arrange(desc(m)) %>% 
#   pivot_wider(-4, 
#               names_from = comb, 
#               values_from = wpd_med)

data_pcp <- data_pcp %>% ungroup()%>% filter(group %in% groupge2$group)
parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 3:ncol(data_pcp),
                   groupColumn = "group",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.3,
                   order = "allClass",
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd") 
# 
# ggparcoord(data_pcp, scale = "globalminmax", columns=3:ncol(data_pcp),groupColumn = "group", alphaLines = 0.8) + geom_boxplot(aes_string(group = "variable"), width = 0.1, alpha = 0.1,
#                  outlier.color = NA) +  scale_color_brewer(palette = "Dark2") 
#+
 # geom_point(data = data, aes(x=comb, y = wpd))

@@

@@

_Regular scale_

```{r}
(p1 + p2)/ (p3 + p4)
(p5 + p6)/(p7 +p8)
```

_Same scale_

```{r}
(p1 +ylim(0,3) + p2 +ylim(0,3))/ (p3+ylim(0,3) + p4+ylim(0,3))
(p5 +ylim(0,3) + p6 +ylim(0,3))/ (p7+ylim(0,3) + p8+ylim(0,3))
```
# PCA based clustering

```{r}
pc <- prcomp(data_pcp[c(-1,-2)], center = F, scale = F)
summary(pc)
# First two principal componentsafc
comp <- data.frame(pc$x[,1:2])

#plot(comp, pch=16, col=rgb(0,0,0,0.5))

hc = stats::hclust(dist(comp))
data_pcp_all <- data_pcp %>% bind_cols(comp)

library(RColorBrewer)
library(scales)
data_pcp_all %>% ggplot(aes(x=PC1, y = PC2)) + geom_point(aes(color = group)) +
  scale_color_brewer(palette = "Dark2") + theme_bw() + theme(legend.position = "bottom")

data_pcp_all %>%
  pivot_longer(-c(1, 2,8, 9), names_to = "variables", values_to = "wpd") %>% 
  ggplot(aes(x=wpd, y = PC1)) + geom_point(aes(color = group)) +
  scale_color_brewer(palette = "Dark2") + theme_bw() +
  facet_wrap(~variables, nrow = 1) + theme(legend.position = "bottom")
```

One dimensional scale since PC1 explains almost all variation?

>>>>

```{r}
ggpairs_nocolor <- ggpairs(data_pcp, columns = 3:7)
ggpairs_nocolor

@@

ggpairs_nocolor <- ggpairs(data_pcp_ge2, columns = 3:7)

ggpairs_color <- ggscatmat(data_pcp_ge2, columns=3:7, color = "group") +
  scale_colour_brewer(palette="Set2")
ggpairs_color

# 
# data_order <- data %>% 
#   group_by(comb) %>% 
#   summarise(m = mean(wpd)) %>% 
#   arrange(desc(m))
#   
# data_median = data %>% 
#   group_by(group, comb) %>% 
#   summarize(wpd_med = median(wpd))
# 
# data_pcp <- data_median %>% 
# mutate(group = as.character(group)) %>% 
#   left_join(data_order) %>% 
#   arrange(desc(m)) %>% 
#   pivot_wider(-4, 
#               names_from = comb, 
#               values_from = wpd_med)

data_pcp <- data_pcp %>% ungroup()%>% filter(group %in% groupge2$group)
parcoord <- GGally::ggparcoord(data_pcp ,
                   columns = 3:ncol(data_pcp),
                   groupColumn = "group",
                   showPoints = TRUE, 
                   title = "(c)",
                   alphaLines = 0.3,
                   order = "allClass",
                   scale = "globalminmax"
) + ggplot2::theme_bw() +
  scale_color_brewer(palette = "Dark2") + 
  ggplot2::theme(
    plot.title = ggplot2::element_text(size=10)
  )+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 10)) +
  theme(legend.position = "bottom") +
  xlab("") +
  ylab("wpd")

c(6, 7, 3, 4, 5)
# 
# ggparcoord(data_pcp, scale = "globalminmax", columns=3:ncol(data_pcp),groupColumn = "group", alphaLines = 0.8) + geom_boxplot(aes_string(group = "variable"), width = 0.1, alpha = 0.1,
#                  outlier.color = NA) +  scale_color_brewer(palette = "Dark2") 
#+
 # geom_point(data = data, aes(x=comb, y = wpd))

@@

(p1 + p2)/ (p3 + p4)
(p5 + p6)/(p7 +p8)

@@
